---
title: "Introduction to traviz"
output: rmarkdown::html_vignette
description: >
  Start here to learn how to use traviz. traviz is a package to meaningfully analyze and visualize trajectory data. This package inherits   from sf and the previous trajectories package aiming to bring a modern approach to analyzing and visualizing trajectory data in R. 
vignette: >
  %\VignetteIndexEntry{traviz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
traviz is a package to meaningfully analyze and visualize trajectories in multiple ways. To showcase some of the features of traviz, we'll use [enviroCar](https://envirocar.org/?lng=en) trajectory data in Muenster, Germany. To collect enviroCar data, the [envirocar-py](https://github.com/enviroCar/envirocar-py) package was used. Keep in mind traviz is not limited to vehicular trajectory data only and can work with all types of trajectory data. 
```{r setup}
library(traviz)
```
## Two approaches to trajectories: point data analysis or general track analysis

traviz offers two approaches to dealing with trajectory data. The trajectory can either be analyzed as a whole geometry with the complete line or the trajectory can be analyzed point by point. Both approaches will be showcased. 

## Dataset: enviroCar sampled data

To explore the basics of working with trajectory data in traviz, we'll use the enviroCar data included in the package. This is included as 'ec.trj'. The dataset consists of:
```{r}
ec <- read.csv("/home/john/Desktop/traviz/data/tracks.csv",header = TRUE, check.names = TRUE)
ec[1,]
```
enviroCar trajectory data contains geographic data with geometry points in sf format, a unique identifier for each track, time stamps, and data measurements at those timestamps. 


